---
title: "Covid91 vaccine study FinalFall2022"
author: "Janette Quezada"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
```{r}
malesdata=filter(FinalData,sex=="M")
```


$H_0$: The treatment has no effect on males.  
$H_A$: The treatment has a significant effect on males.


###Graphical description  

```{r}
barchartGC(~infected + treatment,data=malesdata)
barchartGC(~infected + treatment,data=malesdata, type="percent")

```



According to the bar chart among the men who contracted Covid91 about twice as many males that received the placebo received the Covid91 treatment. The barchart shows that the treatment is effective for males.


##Numerical

```{r}
table1 <- xtabs(~infected + treatment, data=malesdata)
rowPerc(table1)
colPerc(table1)
```

According to the numerical table data, among the males that contracted Covid91 68.44% received the placebo, and 31.56% of males aquired the treatment. The data shows the treatment is effective for males.



##Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


In the Chi Squared test has a p-value of 0.000000000005307 which is less than 0.05 in this case we reject the null hypothesis. In the result the treatment does have an effect on males.
The Fisher's Test has a p-value of 0.000000000002364 which is less than 0.05 meaning we will reject the null hypothesis. In conclusion the treatment does have an effect on males.  
Males who received the placebo are 2.18 times more likely to contract Covid91 than males who received the treatment.


###Discussion and Conclusion



In conclusion the treatment is effective in men.Based on the data above, among the males who contracted Covid91 68.44% received the placebo and 31.56% received the treatment. According to the data, among the males who received the placebo, the odds are that they are 2.18 times more likely to contract Covid91 than males who received the treatment, therefore we reject the null hypothesis.

## Females
```{r}
femalesdata=filter(FinalData,sex=="F")
```





## LGBTQ

```{r}
LGBTQ=filter(FinalData,sex=="gay")
```



## Druggies
```{r}
DrugUser=filter(FinalData,sex=='M')
```




```{r}
DrugUser=filter(FinalData,sex=='F')
```

# Overall Results and Conclusions